---
title: "MICB425 Group 1 Project 2"
author: "Team 1: Albert Chang (), Alison Fong (33399149), Karen Lau (16524143), Yaqian
  Luo (59751503), Jessica Ngo (31837131), Peter (Kiet) Truong (36645133)"
date: "version `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_float:
      collapsed: no
---

####Abstract

####Introduction

####Methods

####Results

#####Environment setup and Data Cleaning
```{r echo=FALSE}
library(tidyverse)
library(cowplot)
```

Loading, parsing (initial data contains data for all genes - we are only interested in narG), renaming data.  We only need taxonomy, abundance, and query data.
```{r message=FALSE, echo=FALSE}
narG.DNA.10m = read_tsv("narG/marker_contig/DNA10m_marker_contig_map.tsv") %>%
    filter(Marker=="narG") %>%
    select(Tax.DNA.10 = Confident_Taxonomy, Abund.DNA.10 = Abundance, Query)

narG.RNA.10m = read_tsv("narG/marker_contig/RNA10m_marker_contig_map.tsv") %>%
    filter(Marker=="narG") %>% 
    select(Tax.RNA.10 = Confident_Taxonomy, Abund.RNA.10 = Abundance, Query)

narG.DNA.100m = read_tsv("narG/marker_contig/DNA100m_marker_contig_map.tsv") %>%
    filter(Marker=="narG") %>% 
    select(Tax.DNA.100 = Confident_Taxonomy, Abund.DNA.100 = Abundance, Query)

narG.RNA.100m = read_tsv("narG/marker_contig/RNA100m_marker_contig_map.tsv") %>%
    filter(Marker=="narG") %>%  
    select(Tax.RNA.100 = Confident_Taxonomy, Abund.RNA.100 = Abundance, Query)

narG.DNA.120m = read_tsv("narG/marker_contig/DNA120m_marker_contig_map.tsv") %>%
    filter(Marker=="narG") %>%  
    select(Tax.DNA.120 = Confident_Taxonomy, Abund.DNA.120 = Abundance, Query)

narG.RNA.120m = read_tsv("narG/marker_contig/RNA120m_marker_contig_map.tsv") %>%
    filter(Marker=="narG") %>%  
    select(Tax.RNA.120 = Confident_Taxonomy, Abund.RNA.120 = Abundance, Query)

narG.DNA.135m = read_tsv("narG/marker_contig/DNA135m_marker_contig_map.tsv") %>%
    filter(Marker=="narG") %>%  
    select(Tax.DNA.135 = Confident_Taxonomy, Abund.DNA.135 = Abundance, Query)

narG.RNA.135m = read_tsv("narG/marker_contig/RNA135m_marker_contig_map.tsv") %>%
    filter(Marker=="narG") %>%  
    select(Tax.RNA.135 = Confident_Taxonomy, Abund.RNA.135 = Abundance, Query)

narG.DNA.150m = read_tsv("narG/marker_contig/DNA150m_marker_contig_map.tsv") %>%
    filter(Marker=="narG") %>%  
    select(Tax.DNA.150 = Confident_Taxonomy, Abund.DNA.150 = Abundance, Query)

narG.RNA.150m = read_tsv("narG/marker_contig/RNA150m_marker_contig_map.tsv") %>% 
    filter(Marker=="narG") %>% 
    select(Tax.RNA.150 = Confident_Taxonomy, Abund.RNA.150 = Abundance, Query)

narG.DNA.165m = read_tsv("narG/marker_contig/DNA165m_marker_contig_map.tsv") %>% 
    filter(Marker=="narG") %>% 
    select(Tax.DNA.165 = Confident_Taxonomy, Abund.DNA.165 = Abundance, Query)

narG.RNA.165m = read_tsv("narG/marker_contig/RNA165m_marker_contig_map.tsv") %>% 
    filter(Marker=="narG") %>% 
    select(Tax.RNA.165 = Confident_Taxonomy, Abund.RNA.165 = Abundance, Query)

narG.DNA.200m = read_tsv("narG/marker_contig/DNA200m_marker_contig_map.tsv") %>% 
    filter(Marker=="narG") %>% 
    select(Tax.DNA.200 = Confident_Taxonomy, Abund.DNA.200 = Abundance, Query)

narG.RNA.200m = read_tsv("narG/marker_contig/RNA200m_marker_contig_map.tsv") %>% 
    filter(Marker=="narG") %>% 
    select(Tax.RNA.200 = Confident_Taxonomy, Abund.RNA.200 = Abundance, Query)
```

Data manipulation into a single data frame
```{r, echo=FALSE}
narG.all = narG.DNA.10m %>% 
# Combine the data frames will full_join to keep all the data
  full_join(narG.RNA.10m, by = "Query") %>%
  full_join(narG.DNA.100m, by = "Query") %>% 
  full_join(narG.RNA.100m, by = "Query") %>%
  full_join(narG.DNA.120m, by = "Query") %>% 
  full_join(narG.RNA.120m, by = "Query") %>%
  full_join(narG.DNA.135m, by = "Query") %>% 
  full_join(narG.RNA.135m, by = "Query") %>%
  full_join(narG.DNA.150m, by = "Query") %>% 
  full_join(narG.RNA.150m, by = "Query") %>%
  full_join(narG.DNA.165m, by = "Query") %>% 
  full_join(narG.RNA.165m, by = "Query") %>% 
  full_join(narG.DNA.200m, by = "Query") %>% 
  full_join(narG.RNA.200m, by = "Query") %>% 
# Create a taxonomy variable aggregating all taxonomy columns so as to fill in any NAs that might occur. !is.na means "is not NA", so the following says that the Taxonomy data should be taken from Tax.RNA.10 if that is not NA, or else take it from Tax.DNA.10 if that is not NA, or else Tax.RNA.200, etc. until if all are NA, give Taxonomy of "unclassified"
  mutate(Taxonomy = ifelse(!is.na(Tax.RNA.10), Tax.RNA.10,
                    ifelse(!is.na(Tax.DNA.10), Tax.DNA.10,
                    ifelse(!is.na(Tax.RNA.100), Tax.RNA.100,
                    ifelse(!is.na(Tax.DNA.100), Tax.DNA.100,
                    ifelse(!is.na(Tax.RNA.120), Tax.RNA.120,
                    ifelse(!is.na(Tax.DNA.120), Tax.DNA.120,
                    ifelse(!is.na(Tax.RNA.135), Tax.RNA.135,
                    ifelse(!is.na(Tax.DNA.135), Tax.DNA.135,
                    ifelse(!is.na(Tax.RNA.150), Tax.RNA.150,
                    ifelse(!is.na(Tax.DNA.150), Tax.DNA.150,
                    ifelse(!is.na(Tax.RNA.165), Tax.RNA.165,
                    ifelse(!is.na(Tax.DNA.165), Tax.DNA.165,
                    ifelse(!is.na(Tax.RNA.200), Tax.RNA.200,
                    ifelse(!is.na(Tax.DNA.200), Tax.DNA.200,
                           "unclassified"))))))))))))))) %>% 
# Get rid of the old Tax variables
  select(-starts_with("Tax.")) %>% 
# Gather all the abundance data into 1 column 
  gather("Key", "Abundance", starts_with("Abund")) %>% 
# Turn the Key into Depth and RNA/DNA variables. We can easily do this because we specifically named these variables with period separation when we loaded in the original data
  separate(Key, c("Key","Type","Depth_m"), by = ".") %>% 
# Remove Key variable now that it only contains "Abund". This also serves to reorder the columns so that the very long Query is at the end.
  select(Depth_m, Type, Abundance, Taxonomy, Query) %>% 
# Make sure R knows depth is numerical since it came from a character variable
  mutate(Depth_m = as.numeric(Depth_m)) %>% 
# Separate Taxonomy into columns so we can plot at different levels
  separate(Taxonomy, into = c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species"), sep="; ")
```

Warning above ignored, because not all queries could be classified down to species level, so these cells are filled as NA, which is fine.

##### Final data table
Table 1. Final data of narG gene 
```{r, echo=FALSE}
narG.all
```

##### 1. How does the DNA abundance of narG gene differ with depth?
```{r, echo=FALSE}
narG.all %>% 
# Filter the DNA data
  filter(Type == "DNA") %>% 
# Change NAs to "unclassified" at the level you want to plot. Here we will do Genus
  mutate(Genus = ifelse(is.na(Phylum), "unclassified", Phylum)) %>% 
  group_by(Depth_m) %>%
  mutate(abund_sum=sum(Abundance, na.rm=TRUE))%>%
  group_by(Depth_m)%>%
  summarize(abund_sum=mean(abund_sum))%>%

  
ggplot(aes(x = "narG", y = Depth_m)) +
# Use the size aesthetic to show abundance
  geom_point(aes(size = abund_sum)) +
# Reverse the why axis so depth increases going down
  scale_y_reverse(lim=c(200,0)) +
  labs(y="Depth (m)")+
  theme_classic() +
  scale_size_continuous(name="Abundance",breaks = c(200,400,600,800)) +
  geom_text(aes(label=abund_sum, hjust=1.5))+
  labs(caption = "Figure 1. Abundance of the narG gene (DNA) at different depths",
       x = "") +
  theme(plot.caption = element_text(hjust=0.5, size=rel(1.0)))
```
 

##### 2. What taxa are responsible for narG? Are they the same with depth? With DNA versus RNA?
```{r, echo=FALSE}
# DNA and RNA abundance with depth
DNA.abund.depth=narG.all %>% 
  filter(Type == "DNA") %>% 
  mutate(Genus = ifelse(is.na(Phylum), "unclassified", Phylum)) %>% 
  group_by(Depth_m) %>%
  mutate(abund_sum=sum(Abundance, na.rm=TRUE))%>%
  group_by(Depth_m)%>%
  summarize(DNA=mean(abund_sum))

RNA.abund.depth=narG.all %>% 
  filter(Type == "RNA") %>% 
  mutate(Genus = ifelse(is.na(Phylum), "unclassified", Phylum)) %>% 
  group_by(Depth_m) %>%
  mutate(abund_sum=sum(Abundance, na.rm=TRUE))%>%
  group_by(Depth_m)%>%
  summarize(RNA=mean(abund_sum))

left_join(DNA.abund.depth,RNA.abund.depth,by="Depth_m")%>%
  gather(Type,abund_sum,DNA:RNA)%>%

# Show both RNA and DNA using an x variable  
ggplot(aes(x = Type, y = Depth_m)) +
  geom_point(aes(size = abund_sum)) +
  labs(y="Depth (m)")+
  scale_y_reverse(lim=c(200,0)) +
  theme_classic()+
  scale_size_continuous(name="Abundance",breaks = c(500,1000,1500,2000,3000)) +
  geom_text(mapping=aes(label=abund_sum),hjust=1.2)+
  labs(caption = "Figure 2. Abundance of the narG gene (DNA vs. RNA) at different depths") +
  theme(plot.caption = element_text(hjust=0.5, size=rel(1.0)))
```

##### 3. What taxa are responsible for narG gene? Are they the same with depth? With DNA versus RNA?
```{r fig.width=8, echo=FALSE}
narG.all %>% 
# Change NAs to "unclassified" at the level you want to plot
  mutate(Phylum = ifelse(is.na(Phylum), "unclassified Bacteria", Phylum)) %>% 
  
ggplot(aes(x = Phylum, y = Depth_m)) +
# Use an ifelse statement to make 0 values into NAs so that they don't show up on the plot
# Use position_dodge to keep points from overlapping
  geom_point(aes(size = ifelse(Abundance == 0, NA, Abundance), color = Type), position = position_dodge(0.5)) +
  scale_y_reverse(lim=c(200,0)) +
  labs(y="Depth (m)")+
  theme_classic() +
# Rename legend
  scale_size_continuous(name = "Abundance") +  
  theme(axis.text.x = element_text(angle = 45, hjust =1, size=rel(1.2)))+
  labs(caption = "Figure 3. Abundance of phyla with narG (DNA vs. RNA) at different depths") +
  theme(plot.caption = element_text(hjust=0.5, size=rel(1.1)))
 
```

Alternative representation with faceted data
```{r fig.width=10, echo=FALSE}
narG.all %>% 
# Change NAs to "unclassified" at the level you want to plot
  mutate(Genus = ifelse(is.na(Phylum), "unclassified Bacteria", Phylum)) %>% 
  
ggplot(aes(x = Genus, y = Depth_m, las=2)) +
  geom_point(aes(size = ifelse(Abundance == 0, NA, Abundance))) +
  scale_y_reverse(lim=c(200,0)) +
  labs(y="Depth (m)")+
  facet_wrap(~Type) +
  theme_classic() +
  scale_size_continuous(name = "Abundance") + 
  theme(axis.text.x = element_text(angle = 45, hjust =1))+
  labs(caption = "Figure 3. Abundance of phyla with narG (DNA vs. RNA) at different depths") +
  theme(plot.caption = element_text(hjust=0.5, size=rel(1.3)))
```

##### 4. How does the abundance of narG gene relate to nitrogen species in Saanich?

Loading in data from project 1 and pulling out the geochemical metadata.
```{r fig.width=7, fig.height=7, echo=FALSE}
load("mothur_phyloseq.RData")

metadata = data.frame(mothur@sam_data)

# DNA/RNA abundance with depth
plot1 = narG.all %>% 
# Change NAs to "unclassified" at the level you want to plot
  mutate(Phylum = ifelse(is.na(Phylum), "unclassified", Phylum)) %>% 
  
ggplot(aes(x = Phylum, y = Depth_m)) +
  geom_point(aes(size = ifelse(Abundance == 0, NA, Abundance), color = Type), position = position_dodge(0.5)) +
  scale_y_reverse(lim=c(200,0)) +
  labs(y = "Depth (m)") +
  theme_classic() +
  scale_size_continuous(name = "Abundance")+ 
  theme(axis.text.x = element_text(angle = 45, hjust =1))

#Then, we make our plots for NO2, "plot2", and NO3, "plot3"

plot2 = metadata %>% 
# Order the data by depth  
  arrange(Depth_m) %>% 
  
ggplot(aes(x = NO2_uM, y = Depth_m)) +
    geom_point() +
    geom_path(aes(group = 1)) +
    scale_y_reverse(lim=c(200,0)) +
  theme_classic() +
  labs(y = "Depth (m)",
       x = "NO2 (uM)")
plot3 = metadata %>% 
# Order the data by depth  
  arrange(Depth_m) %>% 
  
ggplot(aes(x = NO3_uM, y = Depth_m)) +
    geom_point() +
    geom_path(aes(group = 1)) +
    scale_y_reverse(lim=c(200,0)) +
  theme_classic() +
  labs(y = "Depth (m)",
       x = "NO3 (uM)")

plot4 = metadata %>% 
# Order the data by depth  
  arrange(Depth_m) %>% 
  
ggplot(aes(x = NH4_uM, y = Depth_m)) +
    geom_point() +
    geom_path(aes(group = 1)) +
    scale_y_reverse(lim=c(200,0)) +
  theme_classic() +
  labs(y = "Depth (m)",
       x = "NH4 (uM)")

plot5 = metadata %>% 
# Order the data by depth  
  arrange(Depth_m) %>% 
  
ggplot(aes(x = N2O_nM, y = Depth_m)) +
    geom_point() +
    geom_path(aes(group = 1)) +
    scale_y_reverse(lim=c(200,0)) +
  theme_classic() +
  labs(y = "Depth (m)",
       x = "N2O (nM)")

plot_grid(plot4, plot5, plot2, plot3, labels=c("A", "B", "C", "D"), axis="l", nrow=2, ncol=2, rel_widths=c(1/2, 1/2, 1/2,1/2))+
  labs(caption = "Figure 4. Nitrogen species in Saanich across depth. 
A) Ammonium (NH4) concentration across depth   B) Nitrous oxide (N2O) concentration across depth 
C) Nitrite (NO2) concentration across depth     D) Nitrate (NO3) concentration across depth") +
  theme(plot.caption = element_text(hjust=0.5, size=rel(0.75)))
```

####Discussion

####References